---
title: "kagro-covid-19"
author: "Mark Hogue"
date: 
output: html_document
---

## Covid-19 Results by Population, US States
### Covid-19 Data from John Hopkins University
#### - JHU/CCSE repository
### Population from US Census 2017 estimate

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo = FALSE, message = FALSE, warning = FALSE}
library(covid19.analytics)
cases_ts_US <- covid19.data("ts-confirmed-US")
death_ts_US <- covid19.data("ts-deaths-US") 
library(magrittr)
library(dplyr)
library(tidyr)
library(RCurl)
library(ggplot2)
library(tibble)
library(readxl)
library(knitr)
library(kableExtra)
```

#### The most recent day in the dataset is: 

```{r echo = FALSE, message = FALSE, warning = FALSE}
# To just use latest data
# Identify latest date:
cat(names(cases_ts_US[ncol(cases_ts_US)]))
```


```{r echo = FALSE, message = FALSE}

# obtain population data
colNames <- c("state_ter", "census2010", "est_base", paste0("est_",2010:2019))
usPopEst <- read_excel("c:/R_files/nst-est2019-01.xlsx",
                       range="A10:M61",
                       col_names = colNames)[c(1,13)]
usPopEst$state_ter <- gsub("\\.","",usPopEst$state_ter)

# Identify differential cases
latest_cases <- cases_ts_US %>% 
  select(c(1:2,
    c((ncol(.) - 7) : ncol(.))))  #the most recent week

names(latest_cases) <- c(names(cases_ts_US[1:2]), 
          names(cases_ts_US[(ncol(cases_ts_US) - 7):ncol(cases_ts_US)]))               
          
new_cases <- data.frame(latest_cases[, 1:2],
      latest_cases[, 4:10] - latest_cases[, 3:9])  

names(new_cases) <- c(names(latest_cases[1:2]), 
                       names(latest_cases[4:10]))

# most recent day
today_cases <- new_cases %>% 
  group_by(Province_State) %>% 
  select(-c(3:(ncol(.) - 1))) %>% #only the most recent
  select(-1)  #drop "US"
names(today_cases) <- c("state_ter", "latest") 
```
#### Most recent day cases:
#### reverse order of cases per million

```{r echo = FALSE, message = FALSE, warning = FALSE}
case_report <- today_cases %>%  
  summarise(cases = sum(latest)) %>% 
  right_join(., usPopEst) %>% 
  filter(cases > 0) %>% 
  mutate("cases_per_million" = cases / est_2019 * 1e6) %>% 
  arrange(-cases_per_million) 
  
kbl(case_report[c(1, 2, 4)], 
    digits = 1, 
    row.names = TRUE) %>% 
  kable_styling(full_width = FALSE)

```

#### Most recent day deaths:
#### reverse order of deaths per million

```{r echo = FALSE, message = FALSE}

# obtain population data
colNames <- c("state_ter", "census2010", "est_base", paste0("est_",2010:2019))
usPopEst <- read_excel("c:/R_files/nst-est2019-01.xlsx",
                       range="A10:M61",
                       col_names = colNames)[c(1,13)]
usPopEst$state_ter <- gsub("\\.","",usPopEst$state_ter)



# Identify differential deaths
latest_deaths <- death_ts_US %>% 
  select(c(1:2,
    c((ncol(.) - 7) : ncol(.))))  #the most recent week

names(latest_deaths) <- c(names(death_ts_US[1:2]), 
          names(death_ts_US[(ncol(death_ts_US) - 7):ncol(death_ts_US)]))               
          
new_deaths <- data.frame(latest_deaths[, 1:2],
      latest_deaths[, 4:10] - latest_deaths[, 3:9])  

names(new_deaths) <- c(names(latest_deaths[1:2]), 
                       names(latest_deaths[4:10]))

# most recent day
today_deaths <- new_deaths %>% 
  group_by(Province_State) %>% 
  select(-c(3:(ncol(.) - 1))) %>% #only the most recent
  select(-1)  #drop "US"

names(today_deaths) <- c("state_ter", "latest") 

case_report <- today_deaths %>%  
  summarise(deaths = sum(latest)) %>% 
  right_join(., usPopEst) %>% 
  filter(deaths > 0) %>% 
  mutate("deaths_per_million" = deaths / est_2019 * 1e6) %>% 
  arrange(-deaths_per_million) 
  
kbl(case_report[c(1, 2, 4)], 
    digits = 1, 
    row.names = TRUE) %>% 
  kable_styling(full_width = FALSE)

```

# Average for past week

#### Most recent week cases:
#### reverse order of cases per million

```{r echo = FALSE, message = FALSE}

# most recent week
past_week_cases <- new_cases %>% 
   select(-1)  #drop "US"

names(past_week_cases) <- c("state_ter", 
                            paste0("latest", "_", 6:0)) 

case_report <- past_week_cases %>%  
  mutate(case_sum = latest_0 +
                        latest_1 +
                        latest_2 +
                        latest_3 +
                        latest_4 +
                        latest_5 +
                        latest_6) %>% 
  group_by(state_ter) %>% 
  summarise(week_avg = sum(case_sum) / 7) %>% 
  right_join(., usPopEst) %>% 
  filter(week_avg > 0) %>% 
  mutate("cases_per_million" = week_avg / est_2019 * 1e6) %>% 
  arrange(-cases_per_million) 
  
kbl(case_report[c(1, 2, 4)], 
    digits = 1, 
    row.names = TRUE) %>% 
  kable_styling(full_width = FALSE)

```

#### Most recent week deaths:
#### reverse order of deaths per million

```{r echo = FALSE, message = FALSE}

# most recent week
past_week_deaths <- new_deaths %>% 
   select(-1)  #drop "US"

names(past_week_deaths) <- c("state_ter", 
                            paste0("latest", "_", 6:0)) 

death_report <- past_week_deaths %>%  
  mutate(death_sum = latest_0 +
                        latest_1 +
                        latest_2 +
                        latest_3 +
                        latest_4 +
                        latest_5 +
                        latest_6) %>% 
  group_by(state_ter) %>% 
  summarise(week_avg = sum(death_sum) / 7) %>% 
  right_join(., usPopEst) %>% 
  filter(week_avg > 0) %>% 
  mutate("deaths_per_million" = week_avg / est_2019 * 1e6) %>% 
  arrange(-deaths_per_million) 
  
kbl(death_report[c(1, 2, 4)], 
    digits = 1, 
    row.names = TRUE) %>% 
  kable_styling(full_width = FALSE)
```
