---
title: "kagro-covid-19"
author: "Mark Hogue"
date: "null"
output:
  html_document: default
  word_document: default
---

## Covid-19 Results by Population, US States
### Covid-19 Data from John Hopkins University
#### - JHU/CCSE repository
### Population from US Census 2017 estimate

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo = FALSE, message = FALSE, warning = FALSE, include = FALSE}
library(covid19.analytics)
cases_ts_US <- covid19.data("ts-confirmed-US")
death_ts_US <- covid19.data("ts-deaths-US") 
library(magrittr)
library(dplyr)
library(tidyr)
library(RCurl)
library(ggplot2)
library(tibble)
library(readxl)
library(knitr)
```


#### The most recent day in the dataset is

```{r echo = FALSE, message = FALSE, warning = FALSE}
# To just use latest data
# Identify latest date:
cat(names(cases_ts_US[ncol(cases_ts_US)]))
```


```{r echo = FALSE, message = FALSE}

# obtain population data
colNames <- c("st" ,"state_ter", "census2010", "est_base", paste0("est_",2010:2019))
usPopEst <- read_excel("c:/R_files/nst-est2019-01.xlsx",
                       range="A10:N61",
                       col_names = colNames)[c(1, 2, 14)]
usPopEst$state_ter <- gsub("\\.","",usPopEst$state_ter)

# Identify differential cases
latest_cases <- cases_ts_US %>% 
  select(c(1:2,
    c((ncol(.) - 7) : ncol(.))))  #the most recent week

names(latest_cases) <- c(names(cases_ts_US[1:2]), 
          names(cases_ts_US[(ncol(cases_ts_US) - 7):ncol(cases_ts_US)]))               
          
new_cases <- data.frame(latest_cases[, 1:2],
      latest_cases[, 4:10] - latest_cases[, 3:9])  

names(new_cases) <- c(names(latest_cases[1:2]), 
                       names(latest_cases[4:10]))

# most recent day
today_cases <- new_cases %>% 
  group_by(Province_State) %>% 
  select(-c(3:(ncol(.) - 1))) %>% #only the most recent
  select(-1)  #drop "US"
names(today_cases) <- c("state_ter", "latest") 
```
#### Most recent day cases per million population

```{r echo = FALSE, message = FALSE, warning = FALSE, comment = ""}
case_report <- today_cases %>%  
  summarise(cases = sum(latest)) %>% 
  right_join(., usPopEst) %>% 
  filter(cases > 0) %>% 
  mutate("per_1M" = cases / est_2019 * 1e6) %>% 
  arrange(-per_1M) %>% 
  mutate(ord = 1:length(per_1M))

case_report <- case_report %>% 
  select(c(ord, st, cases, per_1M)) %>% 
  mutate(ord = paste0(as.character(ord), " ")) %>% 
  mutate(st = paste0(st, " ")) %>% 
  mutate(cases = paste0(as.character(cases), " ")) %>%
  mutate(per_1M = format(per_1M, digits = 2))

print.data.frame(case_report, row.names = FALSE)  
```

#### Most recent day deaths

```{r echo = FALSE, message = FALSE, warning = FALSE, comment = ""}
# Identify differential deaths
latest_deaths <- death_ts_US %>% 
  select(c(1:2,
    c((ncol(.) - 7) : ncol(.))))  #the most recent week

names(latest_deaths) <- c(names(death_ts_US[1:2]), 
          names(death_ts_US[(ncol(death_ts_US) - 7):ncol(death_ts_US)]))               
          
new_deaths <- data.frame(latest_deaths[, 1:2],
      latest_deaths[, 4:10] - latest_deaths[, 3:9])  

names(new_deaths) <- c(names(latest_deaths[1:2]), 
                       names(latest_deaths[4:10]))

# most recent day
today_deaths <- new_deaths %>% 
  group_by(Province_State) %>% 
  select(-c(3:(ncol(.) - 1))) %>% #only the most recent
  select(-1)  #drop "US"

names(today_deaths) <- c("state_ter", "latest") 

death_report <- today_deaths %>%  
  summarise(deaths = sum(latest)) %>% 
  right_join(., usPopEst) %>% 
  filter(deaths > 0) %>% 
  mutate("per_1M" = deaths / est_2019 * 1e6) %>% 
  arrange(-per_1M) %>% 
  mutate(ord = 1:length(per_1M))

death_report <- death_report %>% 
  select(c(ord, st, deaths, per_1M)) %>% 
  mutate(ord = paste0(as.character(ord), " ")) %>% 
  mutate(st = paste0(st, " ")) %>% 
  mutate(deaths = paste0(as.character(deaths), " ")) %>%
  mutate(per_1M = format(per_1M, digits = 2))

print.data.frame(death_report, row.names = FALSE) 

```

# Average for past week

#### Most recent week average daily cases per million

```{r echo = FALSE, message = FALSE, warning = FALSE, comment = ""}

# most recent week
past_week_cases <- new_cases %>% 
   select(-1)  #drop "US"

names(past_week_cases) <- c("state_ter", 
                            paste0("latest", "_", 6:0)) 

case_report <- past_week_cases %>%  
  mutate(case_sum = latest_0 +
                        latest_1 +
                        latest_2 +
                        latest_3 +
                        latest_4 +
                        latest_5 +
                        latest_6) %>% 
  group_by(state_ter) %>% 
  summarise(week_avg = sum(case_sum) / 7) %>% 
  right_join(., usPopEst) %>% 
  filter(week_avg > 0) %>% 
  mutate("per_1M" = week_avg / est_2019 * 1e6) %>% 
  arrange(-per_1M) %>% 
  mutate(ord = 1:length(per_1M))
  
case_report <- case_report %>% 
  select(c(ord, st, week_avg, per_1M)) %>% 
  mutate(ord = paste0(as.character(ord), " ")) %>% 
  mutate(st = paste0(st, " ")) %>% 
  mutate(week_avg = paste0(as.character(format(week_avg, digits = 2)), " ")) %>%
  mutate(per_1M = format(per_1M, digits = 2))

print.data.frame(case_report, row.names = FALSE) 
```

#### Most recent week average daily deaths per million

```{r echo = FALSE, message = FALSE, warning = FALSE, comment = ""}

# most recent week
past_week_deaths <- new_deaths %>% 
   select(-1)  #drop "US"

names(past_week_deaths) <- c("state_ter", 
                            paste0("latest", "_", 6:0)) 

death_report <- past_week_deaths %>%  
  mutate(death_sum = latest_0 +
                        latest_1 +
                        latest_2 +
                        latest_3 +
                        latest_4 +
                        latest_5 +
                        latest_6) %>% 
  group_by(state_ter) %>% 
  summarise(week_avg = sum(death_sum) / 7) %>% 
  right_join(., usPopEst) %>% 
  filter(week_avg > 0) %>% 
  mutate("per_1M" = week_avg / est_2019 * 1e6) %>% 
  arrange(-per_1M) %>% 
  mutate(ord = 1:length(per_1M))
  
death_report <- death_report %>% 
  select(c(ord, st, week_avg, per_1M)) %>% 
  mutate(ord = paste0(as.character(ord), " ")) %>% 
  mutate(st = paste0(st, " ")) %>% 
  mutate(week_avg = paste0(as.character(format(week_avg, digits = 2)), " ")) %>%
  mutate(per_1M = format(per_1M, digits = 2))

print.data.frame(death_report, row.names = FALSE) 

```
